---
title: "Section 6 Exercises"
output: html_notebook
---

## Exercise 6.0

Load the ``tidyverse`` package and the ``starwars`` data set.

```{r}
library(tidyverse)
starwars <- starwars
```


## Exercise 6.1

Create a histogram of the height variable inside the ``starwars`` data set.

```{r}

```

## Exercise 6.2

Add a title to the previous plot using ``ggtitle()``.

```{r}

```

## Exercise 6.3

Change the label of the x axis in the previous plot to say "height in cm".

```{r}

```

## Exercise 6.4

From the ``starwars`` data, create a scatterplot of height vs. mass. What's the problem?

```{r}

```

Compute the log of mass and re-create the scatterplot.

```{r}

```

Now exclude the outlier (use ``dplyr`` functions) and plot again.

```{r}

```

Make the plot loo nicer (add title, change x and y labels)

```{r}

```

## Exercise 6.5

Use the previous scatterplot and add a smooth line to it.

```{r}

```

Now, use set ``method = "lm"`` argument. What does this do?

```{r}

```

## Exercise 6.6

To the previous plot, add color by gender.

```{r}

```

Now, unnest the films variable (using the ``tidyr`` function ``unnest()``) and use the same plot as before, but now faceted by films.

```{r}

```

## Exercise 6.7

Load the ``data/kicker_match_data.csv`` and re-create one of the plot from the lecture video. 

```{r}

```

Now, try to improve upon it, or try out something new.

```{r}

```

