---
title: "Section 6 Solutions"
output: html_notebook
---

## Exercise 6.0

Load the ``tidyverse`` package and the ``starwars`` data set.

```{r}
library(tidyverse)
starwars <- starwars
```


## Exercise 6.1

Create a histogram of the height variable inside the ``starwars`` data set.

```{r}
starwars %>%
  ggplot(aes(x = height)) +
    geom_histogram()
```

## Exercise 6.2

Add a title to the previous plot using ``ggtitle()``.

```{r}
starwars %>%
  ggplot(aes(x = height)) +
    geom_histogram() +
    ggtitle("Historgram of heights of some Star Wars characters")
```

## Exercise 6.3

Change the label of the x axis in the previous plot to say "height in cm".

```{r}
starwars %>%
  ggplot(aes(x = height)) +
    geom_histogram() +
    ggtitle("Historgram of heights of some Star Wars characters") +
    xlab("height in cm")
```

## Exercise 6.4

From the ``starwars`` data, create a scatterplot of height vs. mass. What's the problem?

```{r}

starwars %>%
  ggplot(aes(x = height, y = mass)) +
  geom_point()
# Outlier!

```

Compute the log of mass and re-create the scatterplot.

```{r}

starwars %>%
  ggplot(aes(x = height, y = log(mass))) +
  geom_point()

```

Now exclude the outlier (use ``dplyr`` functions) and plot again.

```{r}

starwars %>%
  filter(mass != max(mass, na.rm = TRUE)) %>%
  ggplot(aes(x = height, y = log(mass))) +
  geom_point()

```

Make the plot loo nicer (add title, change x and y labels)

```{r}

scatter <- starwars %>%
  filter(mass != max(mass, na.rm = TRUE)) %>%
  ggplot(aes(x = height, y = log(mass))) +
  geom_point() +
  ggtitle("Heights vs. Mass of some Star Wars characters",
          subtitle = "Jabba was excluded...") +
  xlab("height in cm") +
  ylab("natural log of mass in kg")
scatter

```

## Exercise 6.5

Use the previous scatterplot and add a smooth line to it.

```{r}

scatter + geom_smooth()

```

Now, use set ``method = "lm"`` argument. What does this do?

```{r}

# instead of a loess or GAM (non-linear),
# the "lm" method will plot a linear trend
scatter + geom_smooth(method = "lm")

```

## Exercise 6.6

To the previous plot, add color by gender.

```{r}

starwars %>%
  filter(mass != max(mass, na.rm = TRUE)) %>%
  replace_na(list(gender = "none")) %>% # this'll make the plot look much nicer
  ggplot(aes(x = height, y = log(mass), color = gender)) +
    geom_point() +
    ggtitle("Heights vs. Mass of some Star Wars characters",
            subtitle = "Jabba was excluded...") +
    xlab("height in cm") +
    ylab("natural log of mass in kg") + 
    geom_smooth(method = "lm")

```

Now, unnest the films variable (using the ``tidyr`` function ``unnest()``) and use the same plot as before, but now faceted by films.

```{r}

starwars %>%
  filter(mass != max(mass, na.rm = TRUE)) %>%
  replace_na(list(gender = "none")) %>% # this'll make the plot look much nicer
  unnest(films) %>%
  ggplot(aes(x = height, y = log(mass), color = gender)) +
    geom_point() +
    ggtitle("Heights vs. Mass of some Star Wars characters",
            subtitle = "Jabba was excluded...") +
    xlab("height in cm") +
    ylab("natural log of mass in kg") + 
    geom_smooth(method = "lm") +
  facet_wrap(~films)

```

## Exercise 6.7

Load the ``data/kicker_match_data.csv`` and re-create one of the plot from the lecture video. 

```{r}
kicker_data <- read_csv("data/kicker_match_data.csv")

kicker_data %>%
  group_by(matchday) %>%
  mutate(mean_distance = mean(distance_km)) %>%
  filter(team %in% some_teams) %>%
  ggplot(aes(x = matchday)) +
  geom_col(aes(y = distance_km, fill = side)) +
  geom_line(aes(y = mean_distance), linetype = "dashed") +
  ggtitle("Distance run by match day")  +
  ylab("distance in km") +
  coord_cartesian(ylim = c(100,125)) +
  facet_wrap(~team)
```

Now, try to improve upon it, or try out something new.

```{r}

kicker_data %>%
  group_by(matchday, division) %>%
  mutate(mean_distance = mean(distance_km),
         distance_deviation = distance_km - mean_distance) %>%
  filter(team %in% some_teams) %>%
  ggplot(aes(x = matchday)) +
  geom_point(aes(y = distance_km, color = distance_deviation, size = distance_deviation^2)) +
  geom_line(aes(y = mean_distance), linetype = "dashed") +
  ggtitle("Distance run by match day")  +
  ylab("distance in km") +
  scale_color_gradient(low = "red", high = "blue") +
  facet_wrap(~team)
# could be made more pretty...

```

