---
title: "Section 6<br>Plotting data with ggplot2"
author: "Max Mantei"
date: ""
output:
  ioslides_presentation: 
    transition: slower
    toc: yes
    toc_float: yes
    highlight: monochrome
    css: ../custom-css.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What we will cover in this section {.build}

A very (very) brief introduction to ``ggplot2``, one of the most (if not the most) sophisticated tool to plot data.

- Basics of ``ggplot2``
- Histograms and count plots
- Scatterplots and line plots
- Plot more with facets

# Basics of ``ggplot2``

## Basics of ``ggplot2`` {.build}

``ggplot2`` is based on a book called *Grammar of Graphics* (by Leland Wilkinson). It embodies a deep philosophy of visualization.

Usually you start with an empty plot for which you specify ``data`` and so called *"aesthetics"* (or ``aes``). These are the basic outlines of your plot.

You then add layers of points, lines, text, etc. 

You can furthermore specify faceting and scaling/orientation of the coordinate system.

``ggplot2`` is part of the tidyverse.

```{r}
library(tidyverse)
```

# Some examples 

## Histograms {.build .smaller}

```{r message=FALSE}
kicker_data <- read_csv("data/kicker_match_data.csv")
```

```{r fig.align='center', fig.height=4, fig.width=7, message=FALSE}
kicker_data %>% 
  ggplot(aes(x = goals)) +
  geom_histogram()
```

## Bar charts {.build .smaller}

```{r fig.align='center', fig.height=4, fig.width=7, message=FALSE}
kicker_data %>% 
  filter(team == "Bayern München") %>%
  ggplot(aes(y = distance_km, x = matchday, fill = side)) +
  geom_bar(stat = "identity") +
  ggtitle("Distance run by match day", subtitle = "Data for Bayern Munich, season 19/20") +
  ylab("distance in km")
```

## Line plots {.build .smaller}

```{r fig.align='center', fig.height=4, fig.width=7, message=FALSE}
some_teams <- c("Bayern München", "Hertha BSC", "Borussia Dortmund", "Hamburger SV")
kicker_data %>% 
  filter(team %in% some_teams) %>%
  ggplot(aes(y = possession_pct, x = matchday, color = team)) +
  geom_line() +
  ggtitle("Ball possession percentage by match day", subtitle = "Data for season 19/20") +
  ylab("ball possesion %")
```

## Scatterplot {.build .smaller}

```{r fig.align='center', fig.height=4, fig.width=7, message=FALSE}
scatter <- kicker_data %>%
  ggplot(aes(y = shots, x = possession_pct)) +
  geom_point() +
  ggtitle("Ball possession and number of shots", subtitle = "Data for season 19/20") +
  xlab("ball possesion %")
scatter
```

## Scatterplot with smooth {.build .smaller}

```{r fig.align='center', fig.height=4, fig.width=7, message=FALSE}
scatter + geom_smooth()
```

## Faceting {.build .smaller}

```{r fig.align='center', fig.height=4, fig.width=7, message=FALSE}
some_teams <- c("Bayern München", "Hertha BSC", "Borussia Dortmund", "Hamburger SV")
kicker_data %>% 
  filter(team %in% some_teams) %>%
  ggplot(aes(y = shots, x = possession_pct, color = team)) +
  geom_point() +
  ggtitle("Ball possession and number of shots", subtitle = "Data for season 19/20") +
  xlab("ball possesion %") +
  geom_smooth(method = "lm") +
  facet_wrap(~team)
```

## Summary

In this section we had a glimpse into the world of ``ggplot2``...

- Basics of ``ggplot2``
- Histograms and count plots
- Scatterplots and line plots
- Plot more with facets

### You should now do the quiz/challenge for this section!


