---
title: "Section 7 Solutions"
output: html_notebook
---

## Exercise 7.0

Load the ``tidyverse`` package and the ``starwars`` data set.

```{r}
library(tidyverse)
starwars <- starwars
```

## Exercise 7.1

From the ``starwars`` data, create a model matrix, with independent variables mass and gender. Gender is a character variable, so what happens with it?

```{r}

starwars %>% 
  model.matrix(~ mass + gender, data = .)
# missing values will be removed,
# categorical data will be turned
# into dummy variables

```

## Exercise 7.2

Load the ``data/kicker_match_data.csv`` as ``kicker_data``. Run a linear regression with goals as the dependent variable and tackling rate (in %) and pass completion rate (in %) as independent variables. Use the ``summary()`` function to see the results.

```{r}

kicker_data <- read_csv("data/kicker_match_data.csv")

lin_reg1 <- lm(goals ~ tackeling_rate_pct + pass_completion_rate_pct, data = kicker_data)

summary(lin_reg1)

```

Now, instead of additive effect, run the model with an interaction between the two independent variables. Inspect the results.

```{r}

lin_reg2 <- lm(goals ~ tackeling_rate_pct*pass_completion_rate_pct, data = kicker_data)

summary(lin_reg2)

```

## Exercise 7.3

Run the previous regression again, but now as a Poisson GLM using the ``glm()`` function with ``family = poisson()``. 

```{r}

pois_reg1 <- glm(
  goals ~ tackeling_rate_pct*pass_completion_rate_pct, 
  data = kicker_data,
  family = poisson(link = "log")
  )

summary(pois_reg1)

```

Compute the predictions for the Poisson GLM. Do these look reasonable to you (hint: you can plot them with ggplot or use ``summary()`` on the vector of predictions)? 

```{r}

pois_reg1_pred <- predict(pois_reg1)

summary(pois_reg1_pred)

ggplot(data = NULL, # we do not provide a dataset here
       aes(x = pois_reg1_pred)) +
  geom_histogram()

# there are negative values, and
# overall the predictions seem too low

```

Now, use ``predict()`` with argument ``type == "response"``. Check out ``?predict.glm()`` and explain (to yourself) the difference between the two predictions.

```{r}
# The default is to return the linear predictor
# on the scale implied by the link function.
# Use type = "response" to apply the inverse link function...
pois_reg1_pred <- predict(pois_reg1, type = "response")

summary(pois_reg1_pred)
# this makes more sense: predicted expectations
# on the scale of the response variable
```

## Exercise 7.4

Load the ``broom`` package.

```{r}
library(broom)
```

Make a tidy tibble of regression coefficients from the Poisson regression.

```{r}
pois_reg1_coefs <- tidy(pois_reg1)
pois_reg1_coefs
```

Filter this tibble so that there is no ``(Intercept)`` term. Then compute the approximate percentage change as a new variable we briefly discussed in the lecture, now using ```mutate()``. Save the resulting tibble as ``poisson_coefs``.

```{r}
poisson_coefs <- pois_reg1_coefs %>%
  filter(term != "(Intercept)") %>%
  mutate(approx_pct_change = 100 * (exp(estimate) - 1))
```

Now, use ``poisson_coefs`` and put it into ggplot. Create a plot with ``term`` on the y axis and the approx. percentage change in goals estimate on the x axis. Add a layer of points.

```{r}
poisson_coefs %>%
  ggplot(aes(y = term, x = approx_pct_change)) +
  geom_point()
```

Now, add a few more regressors (whichever you see fit) to the model and re-do the plot above.

```{r}

kicker_data %>% glm(
  goals ~ side + shots + distance_km + possession_pct + 
    corners + fouled + offside + tackeling_rate_pct*pass_completion_rate_pct, 
  data = .,
  family = poisson(link = "log")
  ) %>%
  tidy() %>%
  filter(term != "(Intercept)") %>%
  mutate(approx_pct_change = 100 * (exp(estimate) - 1)) %>%
  ggplot(aes(y = term, x = approx_pct_change)) +
    geom_point()

```







