---
title: "Section 5<br>Data transformation in the tidyverse"
author: "Max Mantei"
date: ""
output:
  ioslides_presentation: 
    transition: slower
    toc: yes
    toc_float: yes
    highlight: monochrome
    css: ../custom-css.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What we will cover in this section {.build}

In this section we will explore the ``tidyverse``.

- What’s tidy data?
- What’s in the tidyverse?
- The pipe ``%>%``
- Data transformation with ``dplyr``
- Tidy data with ``tidyr``
- More: Factors, strings, and dates

For more info you can explore https://www.tidyverse.org/.

# Tidyverse

## What’s tidy data? {.build}

Principles of tidy data:

- Every column is a variable
- Every row is an observation
- Every entry is a value

Not everything can be converted to tidy data.

Tidy data is useful, because it is similar to a design matrix used in statistics (especially regression).

## What’s in the tidyverse? {.build .smaller}

The ``tidyverse`` is a collection of packages that are designed to work together:

- ``readr`` and ``haven`` to read in data
- ``tibble`` an enhanced version of the ``data.frame``
- ``dplyr`` lets you work inside and with data sets
- ``tidyr`` lets you bring data into tidy format
- ``ggplot2`` to plot your data
- ``forcats`` for factors, ``stringr`` for strings, ``lubridate`` for dates
- ...and more

```{r eval = FALSE}
install.packages("tidyverse") # to install or update all packages
```

You can use the tidyverse by calling:
```{r}
library(tidyverse)
```

## The pipe ``%>%`` {.build .smaller}

The pipe operator can be awkward at first, but it is actually pretty simple and powerful.

It takes whatever is on the left hand side of it and inputs it to a function on the right hand side *as its first argument*.

```{r}
x <- c(2, 6, 3, 0, -1, 5)
x %>% mean()
```

Usually nested function calls are very hard to understand.

```{r}
y <- c(3, 4, 2, 0, -2, 7)
sqrt(mean((y - x)^2))       # calculate root mean squared error (rmse)
```

With the pipe, those operations become a bit easier to read.

```{r}
(y - x)^2 %>% mean() %>% sqrt()
```

Many functions in the ``tidyverse`` are designed to work with the pipe.

# Data transformation with dplyr

## select {.smaller .build}

The ``select`` function lets you select columns form a tibble.

```{r}
starwars %>%
  head(4)
```

```{r}
starwars %>% 
  select(name, height) %>%
  head(4)
```

## select {.smaller .build}

You can use helper functions to help your select multiple columns, such as ``starts_with()``, ``ends_with()``, ``contains()``.

```{r}
starwars %>% 
  select(name, ends_with("color")) %>%
  head(4)
```
```{r}
starwars %>% 
  select(name, contains("_")) %>%
  head(4)
```

## select {.smaller .build}

You can also use the ``:`` operator to select ranges of columns.

```{r}
starwars %>% 
  select(name:mass) %>%
  head(4)
```

The *exclude* variables, you have to use a minus in front of them (also works with helper functions). You can also use ``-c()`` to exclude groups.

```{r}
starwars %>% 
  select(-name, -contains("_"), -c(films, vehicles, starships)) %>%
  head(4)
```

## filter {.smaller .build}

The ``filter`` function lets you filter observations form a tibble.

```{r}
starwars %>% 
  filter(height < 100)
```

## filter {.smaller .build}

You can use multiple conditions separated with a comma.

```{r}
starwars %>% 
  filter(height < 100, eye_color == "brown")
```

You can use the ``is.na()`` function to filter out missing values (negate with ``!``) and you can use the ``between()`` function to filter for value ranges.

```{r}
starwars %>% 
  filter(is.na(hair_color), between(height, 90, 100))
```

## mutate {.smaller .build}

The ``mutate`` function lets you create new variables inside a tibble.

```{r}
starwars %>% 
  select(name, height) %>%
  mutate(height_in_meters = height/100)
```

## mutate {.smaller .build}

The ``mutate`` function is very useful in combination with the ``if_else()`` function. This lets you perform conditional logic row-by-row inside the data set.

```{r}
starwars %>% 
  select(name, height) %>%
  mutate(height_category = if_else(height < 150, "small", "tall"))
                         # if_else(condition   , TRUE   , FALSE )
```

## arrange {.smaller .build}

The ``arrange`` functions lets you re-order the observations inside a tibble.

```{r}
starwars %>% 
  arrange(height)
```

## arrange {.smaller .build}

You can arrange by descending order using ``desc()``.

```{r}
starwars %>% 
  arrange(desc(height)) %>% 
  head(3)
```
```{r}
starwars %>% 
  select(name, height, mass) %>%
  mutate(round_height = round(height/100)) %>%
  arrange(desc(round_height), desc(mass)) %>% # arranged first by "round_height", then "mass" (both descending)
  head(5)
```

## group_by {.smaller .build}

The ``group_by`` function lets you group the tibble. In a grouped tibble, operations are like ``mutate`` and ``arrange`` are performed by group.

```{r}
starwars %>% 
  select(name, height) %>%
  mutate(height_category = if_else(height < 150, "small", "tall")) %>%
  group_by(height_category) %>% 
  mutate(height_category_mean = mean(height),
         deviation = height - height_category_mean)
```

## summarize {.build}

The ``group_by`` function is particularly useful in combination with the ``summarize`` function. 

```{r}
starwars %>% 
  group_by(gender) %>% 
  summarize(height_mean = mean(height, na.rm = TRUE),
            height_sd = sd(height, na.rm = TRUE))
```

## summarize {.build}

With grouped tibbles you can often use the useful ``n()`` function, which returns the number of observations per group.

```{r}
starwars %>% 
  group_by(gender) %>% 
  summarize(height_mean = mean(height, na.rm = TRUE),
            height_sd = sd(height, na.rm = TRUE),
            group_size = n())
```

## join {.smaller}

We can merge together tibbles in various ways.

To illustrate this, it makes sense to have a look at two small tibbles.

```{r}
band_members
```

```{r}
band_instruments
```
Note, that the two tibbles have one common variable (``name``), but different people in them.

## join {.smaller .build}

The ``left_join()`` keeps all observations of the left hand side (from the pipe operator) tibble.

```{r}
band_members %>% left_join(band_instruments) # band should be complete
```

The ``right_join()`` keeps all observations of the right hand side tibble.

```{r}
band_members %>% right_join(band_instruments) # plays should be complete
```

## join {.smaller .build}

The ``inner_join()`` just keeps observations that are present in **both** tibbles.

```{r}
band_members %>% inner_join(band_instruments) # band and plays should be complete
```

And ``full_join()`` will keep **all** observations.

```{r}
band_members %>% full_join(band_instruments) # both band and plays will be incomplete
```

## join {.smaller .build}

What if the two tibbles do not share a variable with the same name?

```{r}
band_instruments2 # "name" is now "artist"
```
You can use the ``by`` argument to specify the key by which to join.

```{r}
band_members %>% full_join(band_instruments2, by = c("name" = "artist"))
```

The left hand side in the ``by`` argument refers to the key variable in the left hand side tibble and the right hand side to the right hand side tibble.

# Tidy data with tidyr
 
## drop_na {.smaller .build}

The ``starwars`` data set has 87 observations.

```{r}
nrow(starwars)
```

We can exclude all observations with at least one missing value in one of the variables.

```{r}
starwars %>% 
  drop_na()
```

Now, we are left with 6 observations!

## drop_na {.smaller .build}

We can also just drop observations with missing values in specific variables.

```{r}
starwars %>% drop_na(height, mass, gender)
```

There are only 56 observations left. The variables `height`, `mass`, `gender` however, are complete.

## pivot {.smaller .build}

Pivoting lets you transform the data into truly tidy format. 

Sometimes it is not so obvious what *tidy* is.

```{r message=FALSE, warning=FALSE}
kicker_data <- read_csv("data/kicker_match_data.csv") %>%
  select(match_id, side, team, goals, shots)
```

```{r}
kicker_data
```

What's the unit of observation?

## pivot {.smaller .build}

We can make the *match* the unit of observation. 

If we want to preserve all information, this will make the data set wider (more columns, less observations), hence the name of the function ``pivot_wider()``.

```{r}
kicker_data %>% 
  pivot_wider(id_cols = match_id, 
              names_from = side, 
              values_from = team:shots)
```

## pivot {}

There is also the inverse operation ``pivot_longer`` (more observations, less columns).

Pivoting is a really powerful tool, but can be quite challenging.

I can highly recommend checking the vignette for more examples!

```{r, eval = FALSE}
vingette("pivot")
```

# More... factors, strings, dates

## Factors, strings, and dates

The ``tidyverse`` has packages to deal with certain data types.

The ``forcats`` package helps you deal with categorical data encoded in ``factor`` variables. Read the vignette:

```{r, eval = FALSE}
vignette("forcats")
```

The ``stringr`` package deals with character vectors and strings.

```{r, eval = FALSE}
vignette("stringr")
```

Something that we will not cover at all in this course are data and time objects. The ``lubridate`` package has a wide selection of functions for working with times and dates in R.

```{r, eval = FALSE}
vignette("lubridate")
```

## Summary

In this section we covered a few of the main features of the ``tidyverse``.

- What’s the tidyverse?
- The pipe ``%>%``
- Many ``dplyr`` functions
- Some ``tidyr`` functions

For more info you can explore https://www.tidyverse.org/.

A good online resource is the book *R for Data Science* freely available at https://r4ds.had.co.nz/

### You should now do the quiz/challenge for this section!


