---
title: "Section 5 Solutions"
output: html_notebook
---

## Exercise 5.0

If you did not already, download the ``tidyverse`` package. If you are done, load it using ``library()``.

```{r}

library(tidyverse)

```

## Exercise 5.1

Use the pipe operator to pipe ``f`` into the ``paste()`` function, so that the resulting vector is saying
"i is equal to i" 
for every number i in ``f``.
Hint: Use the ``.`` (dot) to input f into two positions of ``paste()``.

```{r}

f <- runif(20)

f %>% paste(., "is equal to", .)

```

## Exercise 5.2

Read-in the ``data/kicker_match_data.csv`` using ``readr``. You can use the drop down menu, or use ``read_csv()`` directly. Save the data set to ``kicker_data``.

```{r}

kicker_data <- read_csv("data/kicker_match_data.csv")

```

## Exercise 5.3

From the ``kicker_data`` select match_id, team, goals, and shots.

```{r}

kicker_data %>% select(match_id, team, goals, shots)

```

## Exercise 5.4

From the ``kicker_data`` select team, and all variables with data on passing statistics.

```{r}

kicker_data %>% select(team, starts_with("pass"))

```

## Exercise 5.5

From the ``kicker_data`` select team, and all variable that are coded as percentages (those are ending with ``_pct``).

```{r}

kicker_data %>% select(team, ends_with("_pct"))

```

## Exercise 5.6

From the ``kicker_data`` filter all matches from division 1. How many observations are in the new dataset? How many matches were played in division 1?

```{r}

kicker_data %>% filter(division == 1)

# output shows there are 612 rows
# we know that there are two observations per match, so
# number of matches is 612/2 = 306

```

## Exercise 5.7

From the ``kicker_data`` filter all observations where the number of goals scored by a team is greater or equal to 5. Which team(s) most frequently scored more than 4 goals?

```{r}

kicker_data %>% 
  filter(goals >= 5) %>% 
  count(team, sort = TRUE)

```

## Exercise 5.8

From the ``kicker_data`` filter all observations where ball possession *and* tackling rate (percentage) was fairly even, i.e. between 49% and 51%.

```{r}

kicker_data %>% 
  filter(
    between(possession_pct, 49, 51), 
    between(tackeling_rate_pct, 49, 51)
  )

```

## Exercise 5.9

With the ``kicker_data`` create a new variable, which holds the average distance run *by player*. To simplify, assume there were always 11 players on the field for each team.

```{r}

kicker_data %>%
  mutate(distance_avg_km = distance_km/11)

```


## Exercise 5.10

Group the ``kicker_data`` by match ID. Calculate the sum of goals by match (use ``mutate()``) and then create a new variable called ``boring``, which has the value TRUE if the sum of goals in a match is 0 and FALSE if not.

```{r}

kicker_data %>% 
  group_by(match_id) %>%
  mutate(
    goals_sum = sum(goals),
    boring = if_else(goals_sum == 0, TRUE, FALSE)
  )

```

## Exercise 5.11

From the ``kicker_data`` filter division 1 and then group by team. Use ``summarize()`` to calculate the sum of goals each division 1 team scored. Arrange the dataset by sum of goals (descending).

```{r}

kicker_data %>% 
  filter(division == 1) %>%
  group_by(team) %>%
  summarise(goals = sum(goals)) %>%
  arrange(desc(goals))

```

## Exercise 5.12

Group the ``kicker_data`` by matchday and division. Use ``mutate()`` to calculate the sum of goals per matchday and division. Then calculate each teams share of this sum (hint: simply divide goals by the new calculated sum).

```{r}

kicker_data %>%
  group_by(matchday, division) %>%
  mutate(goals_total_matchday = sum(goals),
         goals_share = goals/goals_total_matchday)

```

## Exercise 5.13

Use the ``pivot_wider()`` function to make the match the unit of observation (just like in the lecture video).

```{r}

kicker_data %>%
  pivot_wider(
    id_cols = c(match_id, division, matchday),
    names_from = side,
    values_from = team:corners,
    names_sep = "_"
  )

```

