---
title: "Section 1<br>Objects and Data Structures I"
author: "Max Mantei"
date: ""
output:
  ioslides_presentation: 
    transition: slower
    toc: yes
    toc_float: yes
    highlight: monochrome
    css: ../custom-css.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What we will cover in this section {.build}

In this section we will cover the basics of R.

- Numbers and Arithmetics
- Strings and Variables (Assignment)
- Booleans and Relational Operators
- Vectors and Matrices

### Things to note...

Everything in R is an object.

Every object has certain properties and functionality (methods).

Learning these basic functions and how they work is like learning the vocabulary and grammar of a new language.

Just like a real language, you will learn R by using (speaking) it. 

# Numbers and Arithmetics

## Numbers {.build}

When you type in a number into the R console it will just give you back that number.

```{r}
3
```

```{r}
-5
```

```{r}
3.14
```

## Numbers {.build}

Remember that everything in R is an object. 

We can use the ``mode()`` function to find out *what kind* of object it is.

```{r}
mode(3)
```

```{r}
mode(-5)
```

```{r}
mode(3.14)
```

## Numbers {.build}

```{r}
3.14 + 2i
```

```{r}
mode(3.14 + 2i)
```

```{r}
3L
```


```{r}
mode(3L)
```


## Arithmetics {.build}

You can use R as an calculator, since it has all the usual arithmetic operations (and more) built in.

```{r}
5 + 3
```

```{r}
8 - 5
```

```{r}
2 * 3
```

```{r}
13 / 8
```

## Arithmetics {.build}

### Integer division

How often does 3 fit into 5?

```{r}
5 %/% 3
```

### Modulus

What is the remainder of of the (integer) division ``5 %/% 3``?

```{r}
5 %% 3
```

## Arithmetics (operator precedence) {.build}

```{r}
2 + 3 * 4
```

```{r}
(2 + 3) * 4
```

```{r}
3 + 5^2
```

```{r}
(3 + 5)^2
```

## Arithmetics (operator precedence) {.build}

...be careful with exponents!

```{r}
3^2 + 2
```

```{r}
3^(2 + 2)
```

```{r}
3^1/2
```

```{r}
3^(1/2)
```

## More math: functions {.build}

```{r}
3^0.5
```

```{r}
sqrt(3)
```

Other available math functions are (not exhaustive)

``log()``, ``exp()``, ``sin()``, ``asin()``, ``cos()``, ``acos()``, ``tan()``, ``atan()``, ``sign()``, ``abs()``, ``...``

To get to the help file of any function just precede it with an ``?`` and leave the parentheses empty or type ``help("...")`` with the name of the function where the ``...`` are.

```{r eval=FALSE}
?log()
help("log")
```

# Strings and characters

## Strings and characters {.build}

You can define character data by wrapping it with quotes (single ``'``, or double ``"``).

```{r}
"a"
```

```{r}
'a'
```

Strings are arrays of characters, i.e. words or full texts:

```{r}
"Hello"
```

```{r}
"Hello World!"
```

## Strings and characters {.build}

Note that R sees all of those as the same *kind* of object.

```{r}
mode("a")
```

```{r}
mode('a')
```

```{r}
mode("Hello")
```

```{r}
mode("Hello World!")
```

## Strings and characters {.build}

You can save numbers as ``character``, but they lose their ``numeric`` properties.

```{r}
"42"
```

```{r}
mode("42")
```

When you try something like ``"42" + 3`` or ``"42" + "3"`` you will receive an error, since R does not know how to add (or subtract, multiply, divide, etc.) text data. 

## Coersion {.build}

You can coerce some ``character`` objects to ``numeric`` objects using the ``as.numeric()`` function.

```{r}
as.numeric("42")
```

```{r}
as.numeric("42") + 3
```

You can only coerce ``character`` objects which have a reasonable ``numeric`` interpretation, e.g. ``as.numeric("-3.14")``. Something like ``as.numeric("a")`` or ``as.numeric("1a")`` will produce an error.

You can also coerce `numeric` objects to ``character`` objects using ``as.character()``.

```{r}
as.character(42)
```

## Strings and characters: functions {.build}

There are a few functions that you can apply to ``character`` objects.

```{r}
paste("Hello", "World!")
```

```{r}
toupper("Hello World!")
```

```{r}
tolower("Hello World!")
```

## Strings and characters: functions {.build}

```{r}
nchar("Hello World!")
```


```{r}
substring("Hello World!", 7,11)
```

What do you think happens when we apply those functions to objects of type ``numeric``?

```{r}
nchar(1408)
```

```{r}
substring(1408, 2,3)
```

## Variable assignment {.build}

In R you can assign some value (of type ``numeric``, or ``character``, etc.) to a *variable* using the assignment operator ``<-``.

```{r}
a <- "Hello"
b <- "World"
```

Assignment does not return anything, but in RStudio you will see those objects appear in the *Environment* pane (top right by default). Typing the variable into the R console will print its value.

```{r}
a
```

```{r}
mode(a)
```

You can use all functions by simply referring to variables.

## Variable assignment {.build}

This obviously also works with ``numeric`` objects.

```{r}
x <- 3.14
y <- -7
```

```{r}
x^2
```

```{r}
x + sin(y)*log(x) - x/y
```

If you change the assignment to a variable, all subsequent code will be affected by that change.

```{r}
x <- 100
x + sin(y)*log(x) - x/y
```

## Variable assignment {.build}

Valid variable names can only 

- contain upper and lowercase letters, 
- numbers (as long as they are **not** in the beginning of the variable name), 
- dots ``.`` (as long as they are **not** followed by a number), and 
- underscores ``_`` (as long as they are **not** in the beginning of the variable name).

## Variable assignment {}

Example                       | valid?
------------------------------|-------
apple                         | yes
y.1                           | yes
A123                          | yes
.val                          | yes
.1st                          | **no**
my_SUPER_long_variable_name.1 | yes
_this_variable                | **no**
2var_name                     | **no**

# Booleans and Relational Operators

## Booleans and relational/logical operators {.build}

Booleans are named after English mathematician, philosopher and logician [George Boole](https://en.wikipedia.org/wiki/George_Boole). With Booleans we can work with logical operations. In R they can take on the values ``TRUE`` (with abbreviation ``T``) and ``FALSE`` (with abbreviation ``F``).

```{r}
TRUE
```

```{r}
mode(FALSE)
```

```{r}
F
```

## Booleans and relational/logical operators {.build}

Booleans (TRUE and FALSE values) are closely related to relational operations, i.e. comparisons of objects or states. 

In R these are the relational operators:

Operator | Meaning             | Example              | Example Result
---------|---------------------|----------------------|---------------
``>``    | greater than        | ``4 > 7``            | FALSE
`<`      | less than           | ``12 < 20``          | TRUE
`>=`     | greater or equal to | ``8 >= 3``           | TRUE
``<=``   | less or equal to    | ``5 <= 5``           | TRUE
``==``   | equal to            | ``"a" == "b"``       | FALSE
``!=``   | not equal to        | ``"this" != "that"`` | TRUE

## Booleans and relational/logical operators {.build}

Booleans are often used in combination with logical operators.

Operator | Meaning      | Example          | Example Result
---------|--------------|------------------|---------------
``&``    | logical and  | ``FALSE & TRUE`` | FALSE
``|``    | logical or   | ``FALSE | TRUE`` | TRUE
``!``    | logical not  | ``!TRUE``        | FALSE

### Use parentheses.

Relational and logical operators can (and often should) be used with parentheses. 

For example ``!(...)`` negates any truth statement withing the parentheses. 

Parentheses can be used to build  complex truth evaluations.

# Vectors and Matrices

## Vectors {.build}

Have you wondered why there a ``[1]`` in front of all R output we have seen so far?

```{r}
3.14
```

There is a saying that everything in R is a vector.

That's of course not true, but R is *heavily* reliant on vectors.

Vectors are collections of values of same type in a particular order.

You can create vectors in R using the concatenate function ``c()``.

```{r}
c(1, 3.14, -91, 5, -8.15, 0, 1408)
```

## Vectors {.build}

```{r}
my_numbers <- c(1, 1, 2, 3, 5, 8, 15)
my_text <- c("Hello", "there")
```

```{r}
my_numbers
mode(my_numbers)
```

```{r}
my_text
mode(my_text)
```

## Vectors: coersion {.build}

What's happening here?

```{r}
lyrics <- c(1, 2, 3, 4, "tell me that you love me more")
```

```{r}
lyrics
```

A ``character`` vector! R will automatically convert to "sensible" modes:

```{r}
c(-5.2, 7, TRUE, 34, F)
```

```{r}
c(-5.2, 7, TRUE, 34, F, "tree")
```

## Vectors: sequences {.build}

### The colon ``:`` operator

```{r}
1:10
```

```{r}
5:-5
```

```{r}
4.5:9.5
```

```{r}
4.5:9.9
```

## Vectors: sequences {.build}

### The sequence ``seq()`` function

```{r}
seq(from = 5, to = 9, by = 0.7)
```

```{r}
seq(5, -4, -0.3)
```

```{r}
seq(from = 5, to = 9, length.out = 9)
```

## Vectors: sequences {.build}

You can build vectors with repeating sequences using the ``rep()`` function.

```{r}
rep("a", times = 3)
```

```{r}
rep(c("a", "b"), times = 3)
```

```{r}
rep(c("a", "b", "c"), each = 3)
```

## Vectors: access {.build}

The values of vectors are accessed through the brackets ``[]``.

The vector ``month.abb`` contains abbreviated month names.

```{r}
month.abb[4]
```

Using a sequence as index:

```{r}
month.abb[6:8]
```

A negative index will exclude the specified element.

```{r}
month.abb[-11]
```

## Vectors: access {.build}

You can use index vectors to access certain values of vectors.

```{r}
fibonacci <- c(1, 1, 2, 3, 5, 8) # use the first few numbers of the famous fibonacci sequence
```

```{r}
fibonacci[c(3, 1, 6)]
```

```{r}
fibonacci[c(-2, -4)] # or fibonacci[-c(2, 4)]
```

(Note that the ``# or fibonacci[-c(2, 4)]`` is a comment. It will not be evaluated.)

```{r}
fibonacci[c(TRUE, TRUE, FALSE, FALSE, FALSE, TRUE)]
```

Note that the logical index vector must be of same length as the indexed vector.

## Vectors: manipulation {.build}

You can insert and extract elements into/from vectors by indexing and assigning.

```{r}
my_months <- month.abb # copy the month.abb vector
my_months[9] <- "Birthday Month"
my_months
```

```{r}
my_months[6:8] <- "Summer"
my_months
```

```{r}
december_abbreviation <- month.abb[12]
december_abbreviation
```

## Vectors: recycling {.build}

```{r}
things_i_like <- c("music", "books", "food", "swimming")
paste("I like", things_i_like)
```

```{r}
paste("I", c("love", "like"), things_i_like)
```

```{r}
paste("I", c("love", "like", "accept"), things_i_like)
```

Note that note all R functions support this. Some will throw a warning or an error.

Always be aware of recycling: this can have unintended consequences!

## Vectors: math with numeric vectors {.build}

```{r}
x <- c(1, -4, 5, 0)
y <- c(-1, 2, 3, 6)
```

Vector math operations in R are usually elementwise.

```{r}
x + y
```

```{r}
x / y
```

Here recycling also works:

```{r}
g <- c(1, 0)
x * g
```

## Vectors: functions {.build}

The ``length()`` function will return the number of elements in a vector.

```{r}
length(month.abb)
```

You can use this functions to reverse the order of a vector.

```{r}
len_month <- length(month.abb)
month.abb[len_month:1]
```

## Vectors: functions {.build}

There are two ways of (re-)ordering vectors.

The ``order()`` function will return an index vector...

```{r}
months_ordering <- order(month.abb, decreasing = TRUE)
months_ordering
```

...which you can then use to re-order the vector.

```{r}
month.abb[months_ordering]
```

Or, you can simply use the ``sort()`` function.

```{r}
sort(month.abb, decreasing = FALSE)
```

## Vectors: functions {.build}

These functions help you summarize the data in a vector. Recall the short fibonacci sequence from before: ``r fibonacci``.

```{r}
summary(fibonacci)
```

```{r}
fib_tab <- table(fibonacci) # save the resulting table
fib_tab # the table shows the counts of each number in the vector
```

```{r}
prop.table(fib_tab) # the argument to this function is actually a table!
```

## Vectors: functions {.build}

```{r}
x <- c(4, -3, 12, 1, 0, 2)
```

You can also directly compute quantities from numeric vectors.

```{r}
sum(x)
```

```{r}
sum(x)/length(x) # with sum and length we can compute the mean
```

```{r}
mean(x) # there is a function for computing the mean...
```

...and more such as: ``median()``, ``min()``, ``max()``, ``sd()``, ``var``, ``range()``, ...

## Vectors: functions {.build}

```{r}
sin(x) # if necessary, functions are applied element-wise
```

```{r}
log(x)
```

```{r}
x > 0
```

```{r}
log(x[x > 0])
```

## Some more logic with vectors {.build}

With a vector like ``x <- c(4, -3, 12,  1,  0,  2)``, we can check some more logical statements.

```{r}
all(x > 0)
```

```{r}
any(x > 0)
```

```{r}
0 %in% x
```

```{r}
which(x > 0)
```

## Some built-in vectors {.build}

```{r}
month.name
```

```{r}
month.abb
```

```{r}
letters
```

```{r}
LETTERS
```

## A short note about ``factors`` {.build .smaller}

There is a special vector type in R that is called ``factor``. 

A ``factor`` is an extension of a vector used for categorical data.

```{r}
g <- rep(0:1, c(2, 4))
g <- factor(g, levels = 0:1, labels = c("male", "female"))
g
```

```{r}
as.numeric(g)
as.character(g)
```

The main advantage of factors is that R "recognizes" that a factor encodes categories and can thus apply appropriate methods.

Examples are dummy coding in regression models, or visualizing categorical data in a plot.

## Matrices

You can create matrices using the ``matrix()`` function.

```{r}
M <- matrix(1:6, nrow = 3, ncol = 2, byrow = FALSE)
M
```

The first argument specifies the elements in the matrix. (You can use recycling here!)

```{r}
mode(M)
```

```{r}
class(M)
```

## Matrices: math {.build}

Matrices have special properties, such as matrix transpose:

```{r}
t(M)
```

Elementwise (or entry-wise) addition of two matrices.

```{r}
N <- matrix(7:12, nrow = 3, ncol = 2, byrow = TRUE)
M + N
```

## Matrices: math {.build .smaller}

Elementwise multiplication (hadamard product)

```{r}
M * N
```

```{r}
M / N
```

Functions are (usually) applied element-wise.

```{r}
log(M)
```

## Matrices: math {.build .smaller}

Matrix multiplication:

```{r}
t(M) %*% N # necessary to take the transpose!
```

```{r}
M %*% t(N)
```

Matrix inversion $X^{-1}$ is possible using ``solve()``

```{r}
solve(t(M) %*% N) # note, inverse only for square matrices!
```

## Matrices: access {.build}

```{r}
M[2, 1]
```

```{r}
M[2, 1:2]
```

```{r}
M[2, 1:2] <- -M[1, 1:2]
M
```

## Matrices: functions {.build}

```{r}
length(M)
```

When applied to a matrix, the function ``length()`` will return the number of elements.

```{r}
nrow(M)
```

```{r}
ncol(M)
```

```{r}
dim(M) # gives dimension of matrix (ROWS x COLUMNS) in a vector
```

## Matrices: functions {.build .smaller}

You can glue matrices (and vectors) together!

Either column-wise with ``cbind()``:

```{r}
cbind(M, N)
```

...or row-wise with ``rbind()``:

```{r}
rbind(M, N)
```

*Be careful with dimensions and keep recycling in mind!*

## Matrices: functions {.build}

You can compute means of each column (row) of a matrix.

```{r}
colMeans(M)
```

```{r}
rowMeans(M)
```

You can also take the column (row) wise sums of a matrix.

```{r}
colSums(M)
```

```{r}
rowSums(M)
```

# Some notes

## Excursion: modes, types, and classes... {}

Example  | mode (``mode()``)             | type (``typeof()``)              | class (``class()``)
---------|------------------|--------------------|-------------------
``3``    | ``r mode(3)``    | ``r typeof(3)``    | ``r class(3)``
``3.14`` | ``r mode(3.14)`` | ``r typeof(3.14)`` | ``r class(3.14)``
``3L``   | ``r mode(3L)``   | ``r typeof(3L)``   | ``r class(3L)``
``3+1i`` | ``r mode(3+1i)`` | ``r typeof(3+1i)`` | ``r class(3+1i)``
``"a"``  | ``r mode("a")``  | ``r typeof("a")``  | ``r class("a")``
``"Hi!"``| ``r mode("Hi!")``| ``r typeof("Hi!")``| ``r class("Hi!")``
``TRUE`` | ``r mode(TRUE)`` | ``r typeof(TRUE)`` | ``r class(TRUE)``
``F``    | ``r mode(F)``    | ``r typeof(F)``    | ``r class(F)``
``c(1,2)``        | ``r mode(c(1,2))``        | ``r typeof(c(1,2))``        | ``r class(c(1,2))``
``matrix(c(1,2))``| ``r mode(matrix(c(1,2)))``| ``r typeof(matrix(c(1,2)))``| ``r class(matrix(c(1,2)))``


## Excursion: special values {}

Example  | mode             | type               | class
---------|------------------|--------------------|-------------------
``NA``   | ``r mode(NA)``   | ``r typeof(NA)``   | ``r class(NA)``
``NaN``  | ``r mode(NaN)``  | ``r typeof(NaN)``  | ``r class(NaN)``
``NULL`` | ``r mode(NULL)`` | ``r typeof(NULL)`` | ``r class(NULL)``
``Inf``  | ``r mode(Inf)``  | ``r typeof(Inf)``  | ``r class(Inf)``
``NA_character_``| ``r mode(NA_character_)``| ``r typeof(NA_character_)``| ``r class(NA_character_)``
``NA_complex_``  | ``r mode(NA_complex_)``  | ``r typeof(NA_complex_)``  | ``r class(NA_complex_)``
``NA_integer_``  | ``r mode(NA_integer_)``  | ``r typeof(NA_integer_)``  | ``r class(NA_integer_)``
``NA_real_``     | ``r mode(NA_real_)``     | ``r typeof(NA_real_)``     | ``r class(NA_real_)``

## Summary {.build}

We covered the basics of R. 

- Numbers and arithmetics
- Strings and variables (assignment)
- Booleans and relational operators
- Vectors and matrices

You can re-watch any part if something is still unclear. 

You will find a lot of resources online to all the covered topics (just google, really!).

### Do the challenge/quiz for this section before you move on