---
title: "Section 1 Exercises"
output: html_notebook
---

## Exercise 1.1

Compute 42 + log(3.13).

Put the code into the following R chunk. To evaluate the expression click 
*Run current chunk* (the green triangle/play button on the right).

```{r}
42 + log(3.13)
```

The result should be ``[1] 43.14103``.

## Exercise 1.2

Look up the help file for the ``log()`` function (use the ``?``).

```{r}

```

## Exercise 1.3

What is the log (base 3) of 9, i.e. $\log_3(9)$? 
(You can hover the mouse over the LaTex expression between the dollar sign to see LaTex formatting).

```{r}

```

## Exercise 1.4

What is the best way to compute 81 to the power of one half, i.e. $81^{1/2}$?
(You can hover the mouse over the LaTex expression between the dollar sign to see LaTex formatting).

```{r}

```

## Exercise 1.5

How many characters are in this sentence:

``The quick brown fox jumps over the lazy dog``

Hint: Create a character variable with this sentence and then use the ``nchar`` function.

```{r}

```

## Exercise 1.6

Compute the square root of ``x``. You need to convert ``x`` to *numeric* before that.

```{r}
x <- "1522756"




```

## Exercise 1.7

Use the ``is.logical`` function on ``a`` and ``b``. What do the results mean?

```{r}
a <- TRUE
b <- "TRUE"


```

## Exercise 1.8

What happens if you use ``as.numeric`` on variables ``k`` and ``l``?
What happens if you use ``as.character``?

```{r}
k <- TRUE
l <- FALSE
```

## Exercise 1.9

Use ``as.logical`` on the following vector. Explain the results.

```{r}
f <- c(1, 0, 3)



```

## Exercise 1.10

Compute the log of the vector ``g`` and store the result to another variable.
Use the functions ``is.nan()`` ans ``is.na()`` on this new vector.

```{r}
g <- c(2, 6, -1, 0)



```

## Exercise 1.11

Extract all negative values from the vector ``g`` (run the previous chunk if you haven't already).
You can use indexes or a logical operation.

```{r}

```


## Exercise 1.12

Are there any values greater that 2 in the vector ``z``? 
Use a logical operation and the ``any()`` function.

How many values are greater than 2?
Use a logical operation and ``sum()``. (Why does this work?)

```{r}
z <- rnorm(n = 1000, mean = 0, sd = 1)



```

## Exercise 1.13

Create a "stupid" calendar:
It should be a character vector, where the elements are

"1 of Jan", "2 of Jan", ..., "29 of Dec", "30 of Dec"

Assume that every month has 30 days. 
Use ``month.abb`` to get the month abbreviations.
Hint: Use the ``rep`` and ``paste`` functions as needed!

```{r}





```





