---
title: "Section 2 Solutions"
output: html_notebook
---

## Exercise 2.1

Write an input prompt asking you for your name. 
Use the ``readline()`` function and its ``prompt`` argument.

```{r}

# this will not work in the chunk, try it in the console!
my_name <- readline(prompt = "What is your name? ")

```

## Exercise 2.2

Before you do this exercise:
**First, click on the "Preview" drop down, go to "Knit Directory" and select "Project Directory"**
*You should always do that if you want to read in data files from the data folder!*

Read in the file "data/example.txt" using the ``readLines()`` function.
Save the result to a variable. 

```{r}

example <- readLines("data/example.txt")
example

```

## Exercise 2.3

Add a new element, a short text, at the end of the vector you created in the previous exercise.
You can use ``c()`` to concatenate the vector from before with your new string element.
Save the new vector to a file called ``example3.txt`` inside the ``data`` folder.

```{r}

example <- c(example, "And with my additional sentence!")
writeLines(example, "data/example3.txt")
readLines("data/example3.txt")

```

## Exercise 2.4

Read in the ``data/kicker_match_data.csv`` using the drop down menu "Import Dataset"
in the "Environment" pane (default: top right).

## Exercise 2.5

Import the ``maketable.dta`` file in the ``data`` folder using the drop down menu (use "From Stata").


