---
title: "Section 2<br>Data I/O"
author: "Max Mantei"
date: ""
output:
  ioslides_presentation: 
    transition: slower
    toc: yes
    toc_float: yes
    highlight: monochrome
    css: ../custom-css.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What we will cover in this section {.build}

This section will cover:

- Getting user input into R
- Reading and writing ``txt`` files
- Packages in R
- Reading and writing ``csv`` files
- Other file formats

Most of the time real life data is messy, so reading and writing data can be tricky. 

This section can not cover all details but should give you a good start to get data into and out of R.

# Getting user input into R

## User input {.build}

To get input from the user and save it a variable, you can use the ``readline()`` function. ``readline()`` will pause and wait for user input and proceed when the user hits ``Enter``. To save the input, just assign to some variable.

``readline()`` has a ``prompt`` argument where you can specify a message for the prompt.

```{r, eval = FALSE}
user_input <- readline(prompt = "Say something: ")
```

Note that the user input is always ``character``, so you might need to covert to other types using ``as.numeric()``, ``as.integer()``, ``as.logical()``, etc.

This is usually only useful if

- not in interactive mode, or 
- if it is part of a long script, or
- you really need to input data manually.

# Reading and writing ``txt`` files

## Reading ``txt`` files {.smaller .build}

We have a text file (with ``.txt`` extension) saved at ``/intro-r-and-textmining/data/example.txt``. Its content is

```
This is an example
With multiple lines
Of text

```
We can read in ``.txt`` files line-by-line using the ``readLines()`` function. As (first) argument we specify the path and file name of the file we want to open.

```{r}
my_text_file <- readLines("data/example.txt")
my_text_file
```

We can also write text to files using ``writeLines()``. Each element of the input vector will be written on a new line.

```{r}
new_text <- c(my_text_file, "...and some more text!")
writeLines(new_text, "data/example2.txt")
```

Writing functions do usually not return anything. But we can check the result:

```{r}
readLines("data/example2.txt")
```

# Packages in R

## Packages in R {.smaller .build}

So far we have used only R's built-in functions.

R is open source software and many people contribute to it, usually by writing packages that extend the functionality of R.

There is one centralized repository for R packages (CRAN), which makes it easy to install most packages. To install a package, you simply use the ``install.packages()`` function. 

```{r, eval = FALSE}
install.packages("readr") # this installs the readr package
```

Additional packages have to be loaded before you can use them via the ``library()`` function. This has to be done once per session. Conventionally you load all needed packages in the beginning of your script file.

```{r, eval = FALSE}
library(readr) # this loads the readr package
```

Some packages come with *vignettes*. These are often little tutorials demonstrating key functionality of a package. They are often worthwhile to check out and work through.

It is often useful to use the more general help tool ``??...`` to access help pages of packages.

```{r, eval = FALSE}
??readr
```

# Reading and writing ``csv`` files

## Read and write ``csv`` files with readr  {.build}

One of the most comma file formats for structured data is ``csv`` (comma separated values). We can read in ``csv`` using the readr package, which we first have to load.

```{r}
library(readr)
```

Then we can use the function ``read_csv()``.

```{r message=FALSE, eval = FALSE}
kicker_data <- read_csv("data/kicker_match_data.csv")
```

Now the data is stored in memory and you can use it.

To write tables of data to disk as a ``csv`` file you can use ``write_csv()``.

```{r eval = FALSE}
write_csv(my_data, "data/my_new_data.csv")
```

We will learn about data frames (tables of data in R) later.

# Other file formats

## Other file formats {.build}

There are packages to read and write almost all possible data formats in R. Aside from ``readr``, there are also

- ``haven``, to read (SPSS, Stata, SAS)
- ``readxl``, to read in Excel files

...and numerous others.

Usually a quick google search will show you a way (often multiple ways) to open a specific file format.

## Summary

In this section you learned how to

- get user input into R,
- use ``readLines`` and ``writeLines``,
- install and load packages in R ,
- use ``read_csv`` and ``write_csv``,
- import other file formats into R.

### You should now do the quiz/challenge for this section!
