---
title: "Section 4 Solutions"
output: html_notebook
---

## Exercise 4.1

Insert the correct conditions into the if and else if statements (replace ###'s).

```{r}
compare <- function(x, a = 2){
  if(x == a){
    result <- paste(x, "is equal to", a)
  } else if(x > a){
    result <- paste(x, "is greater than", a)
  } else{
    result <- paste(x, "is less than", a)
  }
  return(result)
}

compare(0)
compare(-1)
compare(2)
compare(5)
compare(1, 1)
```

## Exercise 4.2

Insert the correct iterators in the loop (replace ###'s).
Note, that you need to have the ``compare`` function working!

```{r}
compare_vector <- function(x, a = 2){
  N <- length(x)
  result <- vector(mode = "character", length = N)
  for(n in 1:N){
    result[n] <- compare(x[n], a)
  }
  return(result)
}

f <- rpois(15, 2)
compare_vector(f)
```

## Exercise 4.3

Remember the "stupid" calendar you build in the section exercises 1?
Now, build a correct calendar vector and add *st*, *nd*, *rd* or *th* after the days:

"1st of Jan", "2nd of Feb", ..., "30th of Dec", "31st of Dec"

There are multiple ways how you could go about this. You can use ``rep`` and ``paste`` 
and make use of recycling. Or you can use loops and/or functions.

```{r}
# this might be useful, but feel free to ignore this
# if you run this chunk, you can use the days_per_month vector in other chunks...
days_per_month <- c(31, 28, 31, 30, 31, 30, 31, 31, 30, 31, 30, 31)
names(days_per_month) <- month.abb
days_per_month
```

```{r}
# build your calendar here

add_suffix <- function(i){
  if(i %in% c(1,21,31)){
    suff <- "st"
  } else if(i %in% c(2, 22)){
    suff <- "nd"
  } else if(i %in% c(3, 23)){
    suff <- "rd"
  } else{
    suff <- "th"
  }
  return(paste0(i, suff))
}

calendar <- vector(mode = "character", length = sum(days_per_month))

counter <- 1
for(i in 1:12){
  for(j in 1:days_per_month[i]){
    calendar[counter] <- paste(add_suffix(j), "of", names(days_per_month[i]))
    counter <- counter + 1
  }
}

calendar
```

## Exercise 4.4

Build a function that takes in a character vector with a single element.
The input will be a string with multiple words (a sentence).

In your function, reverse *every second* word and return the sentence.

Hint: Look up ``strsplit`` and the argument ``split``. 
You might need to ``unlist`` the result of ``strsplit``.

```{r}

reverse_second <- function(string){
  # split up words (unlist to get vector)
  split <- unlist(strsplit(string, split = " "))
  # iterate over words
  for(s in 1:length(split)){
    # change every second word (using modulus logic)
    if(s %% 2 == 0){
      # split every character, unlist to get vector
      chars <- unlist(strsplit(split[s], split = ""))
      # reverse vector of chars
      chars <- chars[length(chars):1]
      # collapse to string
      split[s] <- paste(chars, collapse = "")
    }
  }
  # return single string by collapsing
  return(paste(split, collapse = " "))
}

```

Check your function:

```{r}
# This should result in:
# "the kciuq brown xof jumps revo the yzal dog"
reverse_second("the quick brown fox jumps over the lazy dog")
```

