---
title: "Section 3 Exercise"
output: html_notebook
---

## Exercise 3.1

Read in the ``data/kicker_match_data.csv``. Give it the name ``kicker`` (on the left in the import menu).
How many rows are in that data frame?
How many columns?

```{r}

```

Print out the first 5 rows of the ``kicker`` data frame using ``head()``.

```{r}

```

Print out the last 5 rows of the ``kicker`` data frame using ``tail()``.

```{r}

```


## Exercise 3.2

What are the column names of the ``kicker`` data frame?

```{r}

```

## Exercise 3.3

Extract the ``team`` column using two different ways!

```{r}

```

## Exercise 3.4

Extract rows 30 to 60 of the ``kicker`` data frame.

```{r}

```

## Exercise 3.5

Use a function to find out about the *structure* of the ``kicker`` data frame.

```{r}

```

Use a function to get summaries of each column in the ``kicker`` data frame.

```{r}

```

## Exercise 3.6

Run the following chunk.

```{r}
my_list <- list(Booleans = c(T, F), Letters = LETTERS, Data = data.frame(num = 7:10, month = month.name[7:10]))
```

Get the names of ``my_list``.

```{r}

```

Whats the length of ``my_list``?

```{r}

```

## Exercise 3.7

Extract the "Letters" vector from the list. What are the *3* different ways to do this?

```{r}

```

## Exercise 3.8

Extract the second row of the data frame (called Data) stored in ``my_list``

```{r}

```





