---
title: "Section 3<br>Objects and Data Structures II"
author: "Max Mantei"
date: ""
output:
  ioslides_presentation: 
    transition: slower
    toc: yes
    toc_float: yes
    highlight: monochrome
    css: ../custom-css.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What we will cover in this section {.build}

In this section we will look at some more data structures.

- Data frames
- Lists

# Data Frames

## Data frames {.smaller .build}

In R data sets are stored in ``data.frames``, which is a collection of vectors of same length but potentially different mode/type.

```{r}
my_df <- data.frame(my_numbers = 1:10, my_letters = letters[1:10])
```

```{r}
head(my_df, 3)
```

```{r}
tail(my_df, 2)
```

```{r}
nrow(my_df)
```

```{r}
ncol(my_df)
```

## Data frames {.smaller .build}

```{r}
names(my_df)
```

```{r}
summary(my_df)
```

```{r}
str(my_df)
```

## Data frames {.smaller .build}

Accessing a data frame is similar to accessing a matrix.

```{r}
my_df$my_letters
```

```{r}
my_df[,2]
```

```{r}
my_df[,"my_letters"]
```

```{r}
my_df$my_letters[1:2]
```

```{r}
my_df[1:2,]
```

## Data frames {.build}

```{r}
my_new_df <- cbind(my_df, my_months = month.abb[1:10])
head(my_new_df, 3)
```

```{r}
my_new_df <- rbind(my_df, c(11, "k"))
tail(my_new_df, 3)
```

# Lists

## Lists {.build}

Lists are the most flexible structure for data in R. A list is basically just a collection of objects of any class.

```{r}
my_list <- list(first = c(T, F), second = LETTERS, third = data.frame(num = 7:10, month = month.name[7:10]))
my_list
```

## Lists {.smaller .build}

```{r}
length(my_list)
```

```{r}
names(my_list)
```

```{r}
my_list$first
```

```{r}
my_list[[3]]
```

```{r}
my_list$third$month
```

## Lists {.build}

Use ``str()`` to find out about the structure of a list.

```{r}
str(my_list)
```

Lists can also store other objects: graphs, statistical models, ...

... and other lists! These are then called nested lists.

Usually lists (and especially nested lists) are not so nice to work with. 

The great flexibility comes at the cost of less obvious functionality.

## Summary

In this section we covered ``data.frame`` objects.

Later we will see an extension of ``data.frame``, the ``tibble`` (coming from *tbl*, which is short for table) and cover data set much more in depth.

We also covered ``list`` objects.

Lists can be seen as a container for data of variable size and type.

### You should now do the quiz/challenge for this section!
